Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?

### 2016-01-28 article/article.Rmd
+ generate replicate prioritisations for inference + fixes to article + switch to ecology letters ref style

<!-- NOTES -->
<!-- 200 word limit for abstract -->
<!-- 6 page limit w/o incurring costs -->

@@

# load .rda files
load('results/results.rda')

# load packages
source('R/analysis/00-initialization.R')

@@

## Introduction
Biodiversity is under threat. Over the last few decades, many species have gone extinct [@r270; @r272]. Today, thousands of species are listed as endangered or near threatened in the International Union for Conservation of Nature (IUCN) red list (\url{www.iucnredlist.org}). In the future, species are expected to face even more threats [eg. climate change; @r236; @r243]. In response to this crisis, the overarching aim of conservation is to maximize the long-term persistence of biodiversity [@r255; @r67].

@@

We investigated the relationship between surrogate-based targets and the genetic variation secured in a prioritization using nonlinear mixed-effects models (NLMMs; using the $\texttt{lme4}$ R package). Two models were fit to the adaptive and neutral genetic variation secured using a negative exponential formulation to estimate the rate of the accumulation of genetic variation ($y = 0 + (0 - 1) \times exp(-r \times x)$; $r$ is the accumulate rate). Both models contained intercept and slope random effects to accommodate variation between species.

@@

>>>>

<!-- NOTES -->
<!-- 150 word limit for abstract -->
<!-- 5000 word length limit -->

@@

# load .rda file
session::restore.session('results/results.rda')

@@

<!--  title page -->
Short running title: Genetic surrogates in conservation \newline
Word count: XXXXX \newline
Number of references: XXXX \newline
Number of figures, table and text boxes: 4 \newline

\newpage

<!-- add line numbers -->
\begin{linenumbers}

## Introduction
<!--  main text -->
Biodiversity is under threat. Over the last few decades, many species have gone extinct [@r270; @r272]. Today, thousands of species are listed as endangered or near threatened in the International Union for Conservation of Nature (IUCN) red list (\url{www.iucnredlist.org}). In the future, species are expected to face even more threats [eg. climate change; @r236; @r243]. In response to this crisis, the overarching aim of conservation is to maximize the long-term persistence of biodiversity [@r255; @r67].

@@

We investigated the relationship between surrogate-based targets and the genetic variation secured in a prioritization using nonlinear mixed-effects models (NLMMs; using the $\texttt{lme4}$ R package). Two models were fit to the adaptive and neutral genetic variation secured using a negative exponential formulation to estimate the rate of the accumulation of genetic variation ($y = 0 + (0 - 1) \times exp(-r \times x)$; $r$ is the accumulate rate). Both models accomodated random effects for the accumulation rate ($r$) parameter.

@@

\end{linenumbers}

### 2016-01-28 article/figures.Rmd
+ generate replicate prioritisations for inference + fixes to article + switch to ecology letters ref style

---
fontsize: 11pt
documentclass: article
bibliography: Endnote_lib.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    includes:
      in_header: [preamble-latex.tex, preamble-latex2.tex]
---

@@

```{r, echo=FALSE, fig.width=6, fig.height=3, fig.cap='Summary of single species prioritizations. Single species prioritizations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species. Violin plots show the performance of prioritizations generated using these three sets of targets. Letters denote significant differences ($P < 0.05$).'}
# prepare data for plotting
single.spp.PDF <- expand.grid(
	Prioritisation=unique(single.spp.SDF$Prioritisation),
	Metric=unique(single.spp.SDF$Metric)
) %>% mutate(
	letters=toupper(cld(single.spp.MCP)$mcletters$Letters),
	letter.height=max(single.spp.SDF$value, na.rm=TRUE)+(diff(range(single.spp.SDF$value, na.rm=TRUE))*0.1)
)
# make plot
suppressWarnings({
	ggplot(aes(x=Metric,y=value,fill=Prioritisation),
		data=single.spp.SDF) +
		geom_violin(position=position_dodge(0.9)) +
		geom_text(aes(x=Metric, y=letter.height, label=letters),
			data=single.spp.PDF, position=position_dodge(0.9)) +
		scale_fill_manual(name='Prioritisation',
			values=c('grey80','grey50','grey20')) +
		ylab('Genetic variation secured (%)') +
		xlab('') +
		theme_classic()
})
```

```{r, echo=FALSE, fig.width=8, fig.height=3, fig.cap='Multi-species prioritizations. Panel (a) shows the prioritization generated for using just amount-based targets. Panel (b) shows the prioritization generated using amount-based and surrogate based targets. Panel (c) shows the prioritization generated using amount-based and genetic-based targets.'}
# download basemap
data(countriesHigh)
countries.FPLY <- countriesHigh[
	countriesHigh$ADMIN %in% c(
		'Italy', 'Switzerland', 'France', 'Austria',
		'Germany', 'Slovenia', 'Croatia', 'Hungary',
		'Monaco', 'Germany', 'Slovakia', 'Czech Republic'
	)
,] %>% spFortify
# prepare data for plotting
multi.spp.grid.FPLY <- grid.PLY
for (i in seq_along(multi.spp.prioritisations))
	multi.spp.grid.FPLY@data[[paste0('v',i)]] <- selections(multi.spp.prioritisations[[i]])
multi.spp.grid.FPLY <- spFortify(multi.spp.grid.FPLY)
# make maps
do.call(
	grid.arrange,
	append(
		llply(
			seq_along(multi.spp.prioritisations),
			function(i) {
				ggplot() +
					geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
						fill='grey20', color='grey80') +
					geom_polygon(data=multi.spp.grid.FPLY, aes_string(x='long', y='lat', 
						group='group', fill=paste0('v',i)),
						alpha=0.8, color='grey10') +
					guides(fill=guide_legend(title=' ')) +
					theme_classic() +
					theme(axis.ticks=element_blank(), axis.text=element_blank(),
						plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(),
						legend.position='none') +
					coord_cartesian(
						xlim=buffered.range(multi.spp.grid.FPLY$long, 0.05),
						ylim=buffered.range(multi.spp.grid.FPLY$lat, 0.05)
					) +
					xlab('') +
					ylab('') +
					annotate('text',
						x=min(multi.spp.grid.FPLY$long)+diff(range(multi.spp.grid.FPLY$long)*0.1),
						y=min(multi.spp.grid.FPLY$lat)+diff(range(multi.spp.grid.FPLY$lat)*1.01),
						label=letters[i], hjust=1, vjust=1, color='white', size=8)
			}
		),
		list(nrow=1)
	)
)
```

```{r, echo=FALSE, fig.width=6, fig.height=3, fig.cap='Summary of multi-species prioritizations. Three prioritizations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species. Data shows the performance of these prioritizations based on how much genetic variation they explain. See Figure 1 caption for conventions.'}
# prepare data for plotting
multi.spp.PDF <- expand.grid(
	Prioritisation=unique(multi.spp.SDF$Prioritisation),
	Metric=unique(multi.spp.SDF$Metric)
) %>% mutate(
	letters=toupper(cld(multi.spp.MCP)$mcletters$Letters),
	letter.height=max(multi.spp.SDF$value, na.rm=TRUE)+(diff(range(multi.spp.SDF$value, na.rm=TRUE))*0.1)
)
# make plot
suppressWarnings({
	ggplot(aes(x=Metric,y=value,fill=Prioritisation),
		data=multi.spp.SDF) +
		geom_violin(position=position_dodge(0.9)) +
		geom_text(aes(x=Metric, y=letter.height,label=letters),
			data=multi.spp.PDF, position=position_dodge(0.9)) +
		scale_fill_manual(name='Prioritisation',
			values=c('grey80','grey50','grey20')) +
		ylab('Genetic variation secured (%)') +
		xlab('') +
		theme_classic()
})
```

```{r, echo=FALSE, fig.width=8, fig.height=3, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations. The left panel shows the relationship between environmental surrogates and adaptive genetic variation, and the right for geographic surrogates and neutral genetic variation. Transparent gray lines denote trneds for each species, and black lines show general trends across all species.'}
### prepare data frames for plotting
## environmental vs adaptive
# random effects
env.correlation.RPDF <- expand.grid(Surrogate.target=seq(min(env.correlation.DF$Surrogate.target), max(env.correlation.DF$Surrogate.target), length.out=100),
	Species=unique(env.correlation.DF$Species))
env.correlation.RPDF$r.ranef <- fixef(env.correlation.NLMM)[[1]] + ranef(env.correlation.NLMM)$Species$r[env.correlation.RPDF$Species]
env.correlation.RPDF$fit <- 1 - (1 - 0) * exp(- env.correlation.RPDF$r.ranef * env.correlation.RPDF$Surrogate.target)
# fixed effects
env.correlation.FPDF <- data.frame(Surrogate.target=seq(min(env.correlation.DF$Surrogate.target), max(env.correlation.DF$Surrogate.target), length.out=100))
env.correlation.FPDF$fit <- 1 - (1 - 0) * exp(- fixef(env.correlation.NLMM)[[1]] * env.correlation.FPDF$Surrogate.target)

>>>>

---
fontsize: 11pt
documentclass: article
bibliography: Endnote_lib.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    keep_tex: yes
    includes:
      in_header: [preamble-latex.tex, preamble-latex2.tex]
---

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=3, fig.cap='Summary of single species prioritizations. Single species prioritizations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species. Violin plots show the performance of prioritizations generated using these three sets of targets. Points indicate the average in each group. Letters denote significant differences ($P < 0.05$).'}
# prepare data for plotting
single.spp.PDF <- expand.grid(
	Prioritisation=unique(single.spp.SDF$Prioritisation),
	Metric=unique(single.spp.SDF$Metric)
) %>% mutate(
	letters=toupper(cld(single.spp.MCP)$mcletters$Letters),
	letter.height=max(single.spp.SDF$value, na.rm=TRUE)+(diff(range(single.spp.SDF$value, na.rm=TRUE))*0.1)
)

# make plot
suppressWarnings({
	ggplot(aes(x=Metric,y=value,fill=Prioritisation),
		data=single.spp.SDF) +
		geom_violin(position=position_dodge(0.9)) +
		geom_text(aes(x=Metric, y=letter.height, label=letters),
			data=single.spp.PDF, position=position_dodge(0.9)) +
		stat_summary(fun.y=mean, geom='point', size=2,
			position=position_dodge(0.9)) +
		scale_fill_manual(name='Prioritisation',
			values=c('grey80','grey50','grey20')) +
		ylab('Genetic variation secured (%)') +
		xlab('') +
		theme_classic()
})
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height=3, fig.cap='Multi-species prioritizations. Panel (a) shows the prioritization generated for using just amount-based targets. Panel (b) shows the prioritization generated using amount-based and surrogate based targets. Panel (c) shows the prioritization generated using amount-based and genetic-based targets.'}
# download basemap
data(countriesHigh)
countries.FPLY <- countriesHigh[
	countriesHigh$ADMIN %in% c(
		'Italy', 'Switzerland', 'France', 'Austria',
		'Germany', 'Slovenia', 'Croatia', 'Hungary',
		'Monaco', 'Germany', 'Slovakia', 'Czech Republic'
	)
,] %>% spFortify
# prepare data for plotting
multi.spp.grid.FPLY <- grid.PLY
for (i in seq_along(multi.spp.prioritisations))
	multi.spp.grid.FPLY@data[[paste0('v',i)]] <- selections(multi.spp.prioritisations[[i]])
multi.spp.grid.FPLY <- spFortify(multi.spp.grid.FPLY)
# make maps
do.call(
	grid.arrange,
	append(
		llply(
			seq_along(multi.spp.prioritisations),
			function(i) {
				ggplot() +
					geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
						fill='grey20', color='grey80') +
					geom_polygon(data=multi.spp.grid.FPLY, aes_string(x='long', y='lat', 
						group='group', fill=paste0('v',i)),
						alpha=0.8, color='grey10') +
					guides(fill=guide_legend(title=' ')) +
					theme_classic() +
					theme(axis.ticks=element_blank(), axis.text=element_blank(),
						plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(),
						legend.position='none') +
					coord_cartesian(
						xlim=buffered.range(multi.spp.grid.FPLY$long, 0.05),
						ylim=buffered.range(multi.spp.grid.FPLY$lat, 0.05)
					) +
					xlab('') +
					ylab('') +
					annotate('text',
						x=min(multi.spp.grid.FPLY$long)+diff(range(multi.spp.grid.FPLY$long)*0.1),
						y=min(multi.spp.grid.FPLY$lat)+diff(range(multi.spp.grid.FPLY$lat)*1.01),
						label=letters[i], hjust=1, vjust=1, color='white', size=8)
			}
		),
		list(nrow=1)
	)
)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=3, fig.cap='Summary of multi-species prioritizations. Three prioritizations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species. Data shows the performance of these prioritizations based on how much genetic variation they explain. See Figure 1 caption for conventions.'}
# prepare data for plotting
multi.spp.PDF <- expand.grid(
	Prioritisation=unique(multi.spp.SDF$Prioritisation),
	Metric=unique(multi.spp.SDF$Metric)
) %>% mutate(
	letters=toupper(cld(multi.spp.MCP)$mcletters$Letters),
	letter.height=max(multi.spp.SDF$value, na.rm=TRUE)+(diff(range(multi.spp.SDF$value, na.rm=TRUE))*0.1)
)
# make plot
suppressWarnings({
	ggplot(aes(x=Metric,y=value,fill=Prioritisation),
		data=multi.spp.SDF) +
		geom_violin(position=position_dodge(0.9)) +
		geom_text(aes(x=Metric, y=letter.height,label=letters),
			data=multi.spp.PDF, position=position_dodge(0.9)) +
		stat_summary(fun.y=mean, geom='point', size=2,
			position=position_dodge(0.9)) +
		scale_fill_manual(name='Prioritisation',
			values=c('grey80','grey50','grey20')) +
		ylab('Genetic variation secured (%)') +
		xlab('') +
		theme_classic()
})
```

```{r, echo=FALSE,  warning=FALSE, message=FALSE, fig.width=8, fig.height=3, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations. The left panel shows the relationship between environmental surrogates and adaptive genetic variation, and the right for geographic surrogates and neutral genetic variation. Transparent gray lines denote trneds for each species, and black lines show general trends across all species.'}
### prepare data frames for plotting
## environmental vs adaptive
# random effects
env.correlation.RPDF <- expand.grid(Surrogate.target=seq(min(env.correlation.DF$Surrogate.target), max(env.correlation.DF$Surrogate.target), length.out=100),
	Species=unique(env.correlation.DF$Species))
env.correlation.RPDF$r.ranef <- fixef(env.correlation.NLMM)[[1]] + ranef(env.correlation.NLMM)$Species$r[env.correlation.RPDF$Species]
env.correlation.RPDF$fit <- 1 - (1 - 0) * exp(- env.correlation.RPDF$r.ranef * env.correlation.RPDF$Surrogate.target)
# fixed effects
env.correlation.FPDF <- data.frame(Surrogate.target=seq(min(env.correlation.DF$Surrogate.target), max(env.correlation.DF$Surrogate.target), length.out=100))
env.correlation.FPDF$fit <- 1 - (1 - 0) * exp(- fixef(env.correlation.NLMM)[[1]] * env.correlation.FPDF$Surrogate.target)
